---
title: "Introduction to gglite"
vignette: >
  %\VignetteIndexEntry{Introduction to gglite}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

The **gglite** package provides a lightweight R interface to the
[AntV G2](https://g2.antv.antgroup.com/) JavaScript visualization library. It
follows the _Grammar of Graphics_ framework---the same theoretical foundation
behind **ggplot2**---but renders interactive, web-based charts powered by G2.

A visualization in gglite is built by composing independent layers:

1. **Data** -- the data frame you want to visualize.
2. **Marks** (geometries) -- the visual shapes representing data (points,
   lines, bars, ...).
3. **Encodings** (aesthetics) -- mappings from data columns to visual channels
   (position, color, size, ...).
4. **Scales** -- control how data values translate to visual values.
5. **Coordinates** -- the coordinate system (Cartesian, polar, ...).
6. **Transforms** -- statistical or layout transforms applied to the data.
7. **Facets** -- split data into multiple panels.
8. **Themes** -- overall visual styling.
9. **Components** -- axes, legends, titles, tooltips, and labels.

Each layer is added with the pipe operator `|>`, so building a chart reads
naturally from left to right.

## Data and encodings

Every chart starts with `g2()`, which accepts a data frame and aesthetic
mappings. Column names are always passed as **character strings** (no
non-standard evaluation):

```{r}
library(gglite)
g2(mtcars, x = 'mpg', y = 'hp') |> mark_point()
```

You can also set encodings later with `encode()`:

```{r}
g2(mtcars) |> encode(x = 'mpg', y = 'hp', color = 'cyl') |> mark_point()
```

## Marks (geometries)

Marks are the visual building blocks. gglite provides 35+ mark types. Here are
the most common ones.

### Points (scatter plot)

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species') |>
  mark_point()
```

### Lines

```{r}
df = data.frame(
  x = rep(1:5, 2), y = c(3, 1, 4, 1, 5, 2, 7, 1, 8, 3),
  group = rep(c('A', 'B'), each = 5)
)
g2(df, x = 'x', y = 'y', color = 'group') |> mark_line()
```

### Bars (intervals)

```{r}
df = data.frame(x = c('A', 'B', 'C', 'D'), y = c(3, 7, 2, 5))
g2(df, x = 'x', y = 'y') |> mark_interval()
```

### Areas

```{r}
df = data.frame(x = 1:10, y = c(3, 1, 4, 1, 5, 9, 2, 6, 5, 3))
g2(df, x = 'x', y = 'y') |> mark_area()
```

### Box plots

```{r}
g2(iris, x = 'Species', y = 'Sepal.Width') |> mark_boxplot()
```

### Combining marks

Multiple marks can be layered on the same chart:

```{r}
df = data.frame(x = c('A', 'B', 'C'), y = c(3, 7, 2))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  mark_text(encode = list(text = 'y'))
```

## Scales

Scales control how data values map to visual properties. Use helpers like
`scale_x()`, `scale_y()`, and `scale_color()` to configure scales:

```{r}
g2(mtcars, x = 'mpg', y = 'hp', color = 'wt') |>
  mark_point() |>
  scale_y(type = 'log') |>
  scale_color(palette = 'viridis')
```

Custom domain and range:

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  scale_x(domain = c(10, 35)) |>
  scale_y(domain = c(0, 400))
```

## Coordinates

Coordinate systems change how positional encodings are interpreted. gglite
supports Cartesian (default), polar, theta, and radial coordinates.

### Polar coordinates (rose chart)

```{r}
df = data.frame(x = c('A', 'B', 'C', 'D'), y = c(3, 7, 2, 5))
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  coord_polar()
```

### Theta coordinates (pie chart)

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  transform_('stackY') |>
  coord_theta(innerRadius = 0.5)
```

### Transposing axes

`coord_transpose()` swaps x and y (similar to ggplot2's `coord_flip()`):

```{r}
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  coord_transpose()
```

## Transforms

Transforms modify the data before rendering. Use `transform_()` to apply
statistical or layout transforms:

### Stacked bars

```{r}
df = data.frame(
  x = rep(c('A', 'B', 'C'), each = 2), y = c(3, 2, 5, 4, 1, 6),
  color = rep(c('a', 'b'), 3)
)
g2(df, x = 'x', y = 'y', color = 'color') |>
  mark_interval() |>
  transform_('stackY')
```

### Dodged bars

```{r}
g2(df, x = 'x', y = 'y', color = 'color') |>
  mark_interval() |>
  transform_('dodgeX')
```

### Stacked area chart

```{r}
df = data.frame(
  x = rep(1:5, 2), y = c(3, 1, 4, 1, 5, 2, 7, 1, 8, 3),
  group = rep(c('A', 'B'), each = 5)
)
g2(df, x = 'x', y = 'y', color = 'group') |>
  mark_area() |>
  transform_('stackY')
```

## Facets

Faceting splits data into panels. Use `facet_rect()` for a grid layout:

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species') |>
  mark_point() |>
  facet_rect(x = 'Species')
```

## Themes

Themes change the overall look. Built-in themes include `theme_classic()`
(default), `theme_classicDark()`, `theme_light()`, `theme_dark()`, and
`theme_academy()`:

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species') |>
  mark_point() |>
  theme_academy()
```

## Components

Components are the non-data elements of a chart: titles, tooltips, axes,
legends, and labels.

### Titles

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  title_('Motor Trend Cars', subtitle = 'mpg vs horsepower')
```

### Tooltips

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  tooltip_(crosshairs = TRUE)
```

### Labels

```{r}
df = data.frame(x = c('A', 'B', 'C', 'D'), y = c(3, 7, 2, 5))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  labels_(text = 'y')
```

## Interactions

Interactions add user-driven behaviors like hovering, brushing, and filtering:

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species') |>
  mark_point() |>
  interact('tooltip') |>
  interact('legendFilter') |>
  interact('brushHighlight')
```

## Putting it all together

Here is a more complete example combining several grammar layers:

```{r}
df = data.frame(
  x = rep(c('Q1', 'Q2', 'Q3', 'Q4'), each = 2),
  y = c(120, 80, 150, 90, 180, 110, 200, 130),
  product = rep(c('Widget', 'Gadget'), 4)
)
g2(df, x = 'x', y = 'y', color = 'product') |>
  mark_interval() |>
  transform_('dodgeX') |>
  scale_color(range = c('#5470c6', '#91cc75')) |>
  title_('Quarterly Sales', subtitle = 'By product line') |>
  interact('tooltip') |>
  interact('elementHighlightByX') |>
  theme_classic()
```
