---
title: Axes and Legends
---

```{r}
library(gglite)
p = g2(mtcars, x = 'mpg', y = 'hp')
```

## Axis Customization

### Axis titles

```{r}
p |>
  axis_x(title = 'Miles per Gallon') |>
  axis_y(title = 'Horsepower')
```

### Hide an axis

```{r}
p |> axis_y(FALSE)
```

### Hide both axes

```{r}
p |>
  axis_x(FALSE) |>
  axis_y(FALSE)
```

### Axis with grid

```{r}
p |>
  axis_x(grid = TRUE) |>
  axis_y(grid = TRUE)
```

### Axis tick count

```{r}
p |>
  axis_x(tickCount = 5) |>
  axis_y(tickCount = 8)
```

## Legend Customization

```{r}
p2 = g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species')
```

### Legend position

```{r}
p2 |> legend_color(position = 'right')
```

### Legend at the top

```{r}
p2 |> legend_color(position = 'top')
```

### Legend at the bottom

```{r}
p2 |> legend_color(position = 'bottom')
```

### Legend at the left

```{r}
p2 |> legend_color(position = 'left')
```

### Hide a legend

```{r}
p2 |> legend_color(FALSE)
```

### Legend with title

```{r}
p2 |> legend_color(position = 'right', title = 'Iris Species')
```

## Dual Y-Axis

Two marks can share the same x-axis but each have their own independent y-axis.
Use `scale_mark_y(independent = TRUE)` on the second mark and
`axis_mark_y(position = 'right')` to place its axis on the right.

```{r}
# Monthly averages of temperature (°F) and wind speed (mph) from airquality
air = aggregate(cbind(Temp, Wind) ~ Month, data = airquality, FUN = mean)
air$Month = month.abb[air$Month]

g2(air, x = 'Month') |>
  mark_interval(encode = list(y = 'Temp')) |>
  style_mark(fill = '#85C5A6', fillOpacity = 0.7) |>
  axis_y(title = 'Temperature (°F)', titleFill = '#85C5A6') |>
  mark_line(encode = list(y = 'Wind')) |>
  style_mark(stroke = 'steelblue', lineWidth = 2) |>
  scale_mark_y(independent = TRUE) |>
  axis_mark_y(
    position = 'right', grid = FALSE,
    title = 'Wind Speed (mph)', titleFill = 'steelblue'
  )
```
