---
title: Canvas Options
---

```{r}
library(gglite)
# This document uses non-default renderers (SVG and WebGL). Declare this
# option globally so all renderers can work on the same page.
options(gglite.renderer = 'svg')
```

## Dimensions

By default a chart auto-fits its container width and uses a height of 480 px.
Use `canvas()` to override:

```{r}
p = g2(mtcars, hp ~ mpg)
p |> canvas(width = 500, height = 300)
```

## Layout Spacing

`padding` is the space between the container edge and the plot area. `margin`
adds space outside the container. `inset` adds space inside the plot area,
pushing axis tick labels inward. Each argument accepts a scalar (all sides) or
`c(top, right, bottom, left)` with `NA` to skip individual sides.

```{r}
p |> canvas(padding = c(0, 150, NA, 100), margin = 10, inset = c(10, 0, 20, -20))
```

## Renderer

G2 ships two bundles with different trade-offs:

- **`g2.min.js`** (~1.1 MB) is a pre-optimized all-in-one bundle that includes
  G2 core, the `@antv/g` graphics engine, and the Canvas renderer. Because
  everything is compiled together it is the smallest and fastest option — but it
  only supports the Canvas renderer.
- **`g2.lite.min.js`** (~976 KB) is the G2 core *without* any graphics engine.
  It must be combined with `@antv/g` and a renderer package (`@antv/g-canvas`,
  `@antv/g-svg`, or `@antv/g-webgl`), which adds weight but gives you a choice
  of renderer.

| `renderer` | Scripts loaded | Approx. total size |
|---|---|---|
| `"Canvas"` (default and optimized) | `g2.min.js` | ~1.1 MB |
| `"Canvas"` (via g2.lite) | `@antv/g` + `@antv/g-canvas` + `g2.lite.min.js` | ~1.5 MB |
| `"SVG"` | `@antv/g` + `@antv/g-svg` + `g2.lite.min.js` | ~1.5 MB |
| `"WebGL"` | `@antv/g` + `@antv/g-webgl` + `g2.lite.min.js` | ~1.9 MB |

SVG output can be inspected element-by-element in browser DevTools and is
therefore well suited to automated DOM-based testing. WebGL is GPU-accelerated
and handles very large numbers of data points with less rendering time.

### Caveat: multi-chart documents

`g2.min.js` and `g2.lite.min.js` **cannot coexist on the same HTML page**.
In a document that contains multiple charts — such as an R Markdown, Quarto,
Shiny, or Jupyter document — setting `options(gglite.renderer)` explicitly
makes sure all plots use `g2.lite.min.js` and avoids loading `g2.min.js`:

```r
options(gglite.renderer = 'svg')   # or 'webgl', 'canvas'
```

Individual charts can still override the renderer via `canvas(renderer = ...)`.
For example, when `options(gglite.renderer = 'svg')` is active, a specific
chart can use `canvas(renderer = 'webgl')` — both use `g2.lite.min.js` so they
do not conflict.

For **standalone plots** previewed in the browser (printed from the console of
an interactive R session), each plot is its own HTML page, so no global option
is needed.

```{r}
p |> canvas(renderer = 'svg')
p |> canvas(renderer = 'webgl')
p |> canvas(renderer = 'canvas')
```

## Extra Options

The `...` argument in `canvas()` passes additional top-level options to
`chart.options()` in JavaScript. For example, `clip` controls whether marks
that extend beyond the plot area are clipped:

```{r}
# clip = FALSE (default): marks can overflow the plot area
p |> scale_y(domain = c(80, 220)) |> canvas(clip = FALSE)
# clip = TRUE: marks are clipped to the plot area
p |> scale_y(domain = c(80, 220)) |> canvas(clip = TRUE)
```

## WebGL for Large Datasets

WebGL is GPU-accelerated and can render far more data points than Canvas or SVG
before frame time degrades. The example below uses 200,000 flight records
(distance vs. departure delay) to illustrate this.

```r
g2("https://vega.github.io/vega-datasets/data/flights-200k.json") |>
  encode(delay ~ distance) |>
  mark_point() |>
  style_mark(size = 2, fillOpacity = 0.2) |>
  canvas(renderer = 'webgl')
```

However, it wouldn't render on my M1 macOS and could even freeze my browser, and
I don't know why...
