---
title: Coordinates
---

```{r}
library(gglite)
```

Coordinate systems control how the positional channels (x and y) are
interpreted. Use `coordinate()` to set the type and options.

## Cartesian (default)

The default coordinate system. Typically you don't need to specify it
explicitly, but you can do so for clarity.

```{r}
df = data.frame(x = c('A', 'B', 'C', 'D'), y = c(3, 7, 2, 5))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  coordinate('cartesian')
```

## Polar

Maps x to angle and y to radius. Useful for rose charts and radar-like
displays.

```{r}
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  coordinate('polar')
```

### Polar with innerRadius (donut rose)

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  coordinate('polar', innerRadius = 0.4)
```

### Polar with custom start/end angles

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  coordinate('polar', startAngle = -pi / 2, endAngle = pi / 2)
```

## Theta

Maps data values to angular extent. Used for pie and donut charts.

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  transform_of('stackY') |>
  coordinate('theta')
```

### Donut chart (theta with innerRadius)

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  transform_of('stackY') |>
  coordinate('theta', innerRadius = 0.5)
```

## Radial

Maps y to the radial direction. Suitable for radial bar charts.

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  coordinate('radial')
```

### Radial with innerRadius

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  coordinate('radial', innerRadius = 0.3)
```

## Transpose (coord_flip)

Swap x and y axes, equivalent to ggplot2's `coord_flip()`. This is a
coordinate transform, not a type.

```{r}
df = data.frame(x = c('A', 'B', 'C', 'D'), y = c(3, 7, 2, 5))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  coord_transpose()
```

### Transpose with polar

```{r}
g2(df, x = 'x', y = 'y', color = 'x') |>
  mark_interval() |>
  coordinate('polar') |>
  coord_transpose()
```
