---
title: Interactions
---

```{r}
library(gglite)
```

Interactions enable user-driven behaviors such as tooltips, brushing, and
highlighting. Use `interact()` to add them.

## Tooltip

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  interact('tooltip')
```

## Element Highlight

Highlight individual elements on hover.

```{r}
g2(mtcars, x = 'mpg', y = 'hp', color = 'cyl') |>
  mark_point() |>
  interact('elementHighlight')
```

## Element Highlight by X

Highlight all elements sharing the same x value.

```{r}
df = data.frame(
  x = rep(c('A', 'B', 'C'), each = 2), y = c(3, 2, 5, 4, 1, 6),
  color = rep(c('a', 'b'), 3)
)
g2(df, x = 'x', y = 'y', color = 'color') |>
  mark_interval() |>
  transform_('dodgeX') |>
  interact('elementHighlightByX')
```

## Element Highlight by Color

```{r}
g2(df, x = 'x', y = 'y', color = 'color') |>
  mark_interval() |>
  transform_('dodgeX') |>
  interact('elementHighlightByColor')
```

## Element Select

Click to select individual elements.

```{r}
g2(mtcars, x = 'mpg', y = 'hp', color = 'cyl') |>
  mark_point() |>
  interact('elementSelect')
```

## Brush Highlight

Drag to brush and highlight a rectangular region.

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  interact('brushHighlight')
```

## Brush X Highlight

Brush only along the x axis.

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  interact('brushXHighlight')
```

## Brush Y Highlight

Brush only along the y axis.

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  interact('brushYHighlight')
```

## Brush Filter

Drag to filter the visible data.

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  interact('brushFilter')
```

## Legend Filter

Click legend entries to filter data.

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species') |>
  mark_point() |>
  interact('legendFilter')
```

## Legend Highlight

Hover over legend entries to highlight.

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species') |>
  mark_point() |>
  interact('legendHighlight')
```

## Multiple Interactions

You can combine several interactions.

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', color = 'Species') |>
  mark_point() |>
  interact('tooltip') |>
  interact('legendFilter') |>
  interact('brushHighlight')
```
