---
title: Labels and Styles
---

```{r}
library(gglite)
```

## Labels

Labels add text annotations to marks. Use `labels_of()` to configure them.

### Basic labels on bars

```{r}
df = data.frame(x = c('A', 'B', 'C'), y = c(3, 7, 2))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  labels_of(text = 'y')
```

### Label position: inside

```{r}
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  labels_of(text = 'y', position = 'inside')
```

### Label position: top

```{r}
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  labels_of(text = 'y', position = 'top')
```

### Labels on a scatter plot

```{r}
df = data.frame(x = 1:5, y = c(3, 7, 2, 9, 5), label = c('a', 'b', 'c', 'd', 'e'))
g2(df, x = 'x', y = 'y') |>
  mark_point() |>
  labels_of(text = 'label')
```

### Multiple label layers

```{r}
df = data.frame(x = c('A', 'B', 'C'), y = c(3, 7, 2))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  labels_of(text = 'x', position = 'inside') |>
  labels_of(text = 'y', position = 'top')
```

## Mark Styles

Use `style_mark()` to set visual properties on the most recently added mark.

### Custom fill and stroke on points

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  style_mark(fill = 'steelblue', stroke = 'white', lineWidth = 1)
```

### Custom fill on bars

```{r}
df = data.frame(x = c('A', 'B', 'C'), y = c(3, 7, 2))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  style_mark(fill = 'coral', stroke = '#333', lineWidth = 0.5)
```

### Semi-transparent area

```{r}
df = data.frame(x = 1:10, y = c(3, 1, 4, 1, 5, 9, 2, 6, 5, 3))
g2(df, x = 'x', y = 'y') |>
  mark_area() |>
  style_mark(fill = 'steelblue', fillOpacity = 0.3)
```

### Line with custom stroke

```{r}
df = data.frame(x = 1:10, y = cumsum(rnorm(10)))
g2(df, x = 'x', y = 'y') |>
  mark_line() |>
  style_mark(stroke = 'tomato', lineWidth = 3)
```

### Dashed line

```{r}
g2(df, x = 'x', y = 'y') |>
  mark_line() |>
  style_mark(stroke = 'steelblue', lineDash = c(6, 3))
```
