---
title: Marks (Geometry Layers)
---

```{r}
library(gglite)
```

## Point Mark (Scatter Plot)

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |> mark_point()
```

### Color by a variable

```{r}
g2(mtcars, x = 'mpg', y = 'hp', color = 'cyl') |> mark_point()
```

### Size by a variable

```{r}
g2(mtcars, x = 'mpg', y = 'hp', size = 'wt') |> mark_point()
```

### Shape by a variable

```{r}
g2(iris, x = 'Sepal.Width', y = 'Sepal.Length', shape = 'Species') |>
  mark_point()
```

### Custom style

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point(style = list(fill = 'steelblue', stroke = 'white', lineWidth = 1))
```

## Line Mark

```{r}
df = data.frame(x = 1:10, y = cumsum(rnorm(10)))
g2(df, x = 'x', y = 'y') |> mark_line()
```

### Multi-series line

```{r}
df = data.frame(
  x = rep(1:5, 2), y = c(3, 1, 4, 1, 5, 2, 7, 1, 8, 3),
  group = rep(c('A', 'B'), each = 5)
)
g2(df, x = 'x', y = 'y', color = 'group') |> mark_line()
```

## Interval Mark (Bar Chart)

```{r}
df = data.frame(x = c('A', 'B', 'C', 'D'), y = c(3, 7, 2, 5))
g2(df, x = 'x', y = 'y') |> mark_interval()
```

### Stacked bar

```{r}
df = data.frame(
  x = rep(c('A', 'B', 'C'), each = 2), y = c(3, 2, 5, 4, 1, 6),
  color = rep(c('a', 'b'), 3)
)
g2(df, x = 'x', y = 'y', color = 'color') |>
  mark_interval() |> transform_of('stackY')
```

### Grouped (dodged) bar

```{r}
g2(df, x = 'x', y = 'y', color = 'color') |>
  mark_interval() |> transform_of('dodgeX')
```

## Area Mark

```{r}
df = data.frame(x = 1:10, y = c(3, 1, 4, 1, 5, 9, 2, 6, 5, 3))
g2(df, x = 'x', y = 'y') |> mark_area()
```

### Stacked area

```{r}
df = data.frame(
  x = rep(1:5, 2), y = c(3, 1, 4, 1, 5, 2, 7, 1, 8, 3),
  group = rep(c('A', 'B'), each = 5)
)
g2(df, x = 'x', y = 'y', color = 'group') |>
  mark_area() |> transform_of('stackY')
```

## Rect Mark

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_rect(
    transform = list(list(type = 'bin', thresholdsX = 10, thresholdsY = 10))
  )
```

## Cell Mark

```{r}
df = expand.grid(x = LETTERS[1:5], y = LETTERS[1:5])
df$value = seq_len(nrow(df))
g2(df, x = 'x', y = 'y', color = 'value') |> mark_cell()
```

## Text Mark

```{r}
df = data.frame(x = c('A', 'B', 'C'), y = c(3, 7, 2))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  mark_text(encode = list(text = 'y'))
```

## Path Mark

```{r}
n = 100
t = seq(0, 4 * pi, length.out = n)
df = data.frame(x = t * cos(t), y = t * sin(t))
g2(df, x = 'x', y = 'y') |> mark_path()
```

## Polygon Mark

```{r}
df = data.frame(x = c(0, 1, 0.5), y = c(0, 0, 1))
g2(df, x = 'x', y = 'y') |> mark_polygon()
```

## Link Mark

```{r}
df = data.frame(x = c(0, 1), y = c(0, 0), x1 = c(1, 2), y1 = c(1, 1))
g2(df) |>
  mark_link(encode = list(x = c('x', 'x1'), y = c('y', 'y1')))
```

## Reference Lines

### Vertical reference line (lineX)

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  mark_line_x(
    data = list(list(x = 20)),
    style = list(stroke = 'red', lineDash = c(4, 4))
  )
```

### Horizontal reference line (lineY)

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  mark_line_y(
    data = list(list(y = 150)),
    style = list(stroke = 'red', lineDash = c(4, 4))
  )
```

## Range Marks

### Full range

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  mark_range(
    data = list(list(x = c(15, 25), y = c(100, 200))),
    encode = list(x = 'x', y = 'y'),
    style = list(fill = 'steelblue', fillOpacity = 0.15)
  )
```

### Horizontal range (rangeX)

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  mark_range_x(
    data = list(list(x = c(15, 25))),
    encode = list(x = 'x'),
    style = list(fill = 'steelblue', fillOpacity = 0.15)
  )
```

### Vertical range (rangeY)

```{r}
g2(mtcars, x = 'mpg', y = 'hp') |>
  mark_point() |>
  mark_range_y(
    data = list(list(y = c(100, 200))),
    encode = list(y = 'y'),
    style = list(fill = 'orange', fillOpacity = 0.15)
  )
```

## Connector Mark

```{r}
df = data.frame(x = c('A', 'B'), y = c(3, 7))
g2(df, x = 'x', y = 'y') |>
  mark_interval() |>
  mark_connector(
    data = list(list(x = 'A', x1 = 'B')),
    encode = list(x = 'x', x1 = 'x1'),
    labels = list(list(text = '+133%'))
  )
```

## Box Plot Mark

```{r}
g2(iris, x = 'Species', y = 'Sepal.Width') |> mark_boxplot()
```

## Liquid Mark

```{r}
g2() |>
  mark_liquid(
    data = list(list(value = 0.6)),
    encode = list(y = 'value'),
    style = list(textContent = '60%')
  )
```
