---
title: Tooltips
---

Tooltips in gglite have two distinct configuration points that match G2's
internal structure:

- **Chart-level** (`tooltip_()` → `interaction.tooltip`): controls tooltip
  *behavior* — crosshairs, shared mode, marker style, position, and rendering.
- **Mark-level** (`mark_*(tooltip = list(...))` → mark spec `tooltip`): controls
  what *data* is shown for a specific mark — which channels to display, value
  formatting, custom item names.

## Chart-Level Tooltip (`tooltip_()`)

All arguments to `tooltip_()` are passed to G2's `interaction.tooltip`. These
options apply across all marks in the chart.

| Option | Default | Description |
|--------|---------|-------------|
| `FALSE` | — | Disable tooltip entirely |
| `crosshairs` | `FALSE` | Show crosshairs on both axes (series marks only) |
| `crosshairsX` | `FALSE` | Show horizontal crosshair only (series marks only) |
| `crosshairsY` | `FALSE` | Show vertical crosshair only (series marks only) |
| `crosshairsStroke` | `'#1b1e23'` | Crosshair line color |
| `crosshairsLineWidth` | `1` | Crosshair line width in pixels |
| `crosshairsLineDash` | `NULL` | Dash pattern, e.g. `c(4, 4)` |
| `crosshairsStrokeOpacity` | `0.5` | Crosshair line opacity (0–1). Set to `1` for a fully opaque line |
| `shared` | `FALSE` | Show all mark values at the same x position |
| `groupName` | auto | Show series/group name prefix in tooltip items |
| `marker` | `TRUE` | Show the hover marker dot |
| `markerType` | `'default'` | Marker shape; `'hollow'` gives an unfilled circle |
| `markerFill` | auto | Marker fill color |
| `markerStroke` | auto | Marker stroke color |
| `markerSize` | auto | Marker size in pixels |
| `body` | `TRUE` | Show tooltip box; `FALSE` shows only the marker |
| `position` | `'right-bottom'` | Tooltip position: `'top'`, `'bottom'`, `'left'`, `'right'`, or a corner such as `'top-left'` |
| `offset` | `c(10, 10)` | x/y pixel offset from the cursor |
| `enterable` | `FALSE` | Allow the cursor to move inside the tooltip box |

### Disable Tooltip

Pass `FALSE` to turn off the tooltip entirely.

```{r}
library(gglite)
g2(mtcars, hp ~ mpg) |> tooltip_(FALSE)
```

### Crosshairs

Crosshairs are reference lines that track the cursor along the x and/or y axes.
They only work with the *series* tooltip path used by line and area marks.

```{r}
air_monthly = aggregate(Temp ~ Month, data = airquality, FUN = mean)
air_monthly$Month = month.abb[air_monthly$Month]
p = g2(air_monthly, Temp ~ Month) |> mark_line() |> mark_point()
p |> tooltip_(crosshairs = TRUE)
```

### Crosshairs on One Axis Only

`crosshairsX = TRUE` shows only the horizontal line; `crosshairsY = TRUE` shows
only the vertical line.

```{r}
p |> tooltip_(crosshairsX = TRUE, crosshairsY = FALSE)
```

```{r}
p |> tooltip_(crosshairsX = FALSE, crosshairsY = TRUE)
```

### Crosshairs Style

Use `crosshairsStroke`, `crosshairsLineWidth`, `crosshairsLineDash`, and
`crosshairsStrokeOpacity` to customize crosshair appearance.

```{r}
p |> tooltip_(
  crosshairs = TRUE, crosshairsStroke = 'tomato', crosshairsLineWidth = 2
)
```

### Default Crosshairs in Line Charts

Series marks (`mark_line()`, `mark_area()`) show a **vertical crosshair by
default** without any `tooltip_()` call. This is controlled by G2's built-in
`crosshairs: true` mark option. If you do not want crosshairs, pass
`crosshairsY = FALSE` (or `crosshairs = FALSE`) to `tooltip_()`:

```{r}
p |> tooltip_(crosshairsY = FALSE)
```

### Marker Style

The hover marker is the dot that snaps to the data point. Use `marker = FALSE`
to hide it, or customize with `markerType`, `markerFill`, `markerStroke`, and
`markerSize`. These examples use a line-only chart so the marker effect is
clearly visible.

```{r}
p_line = g2(air_monthly, Temp ~ Month) |> mark_line()
p_line |> tooltip_(marker = FALSE)
```

```{r}
p_line |> tooltip_(markerType = 'hollow', markerSize = 12)
```

### Hide Tooltip Body

`body = FALSE` shows only the hover marker with no tooltip box.

```{r}
p |> tooltip_(body = FALSE)
```

### Tooltip Position

Control where the tooltip appears relative to the cursor.

```{r}
p |> tooltip_(position = 'top-left')
```

### Tooltip Offset

Adjust the tooltip's x/y offset from the cursor in pixels.

```{r}
p |> tooltip_(offset = c(30, 30))
```

### Enterable Tooltip

`enterable = TRUE` lets the user move the cursor inside the tooltip box,
useful when the tooltip contains interactive content.

```{r}
p |> tooltip_(enterable = TRUE)
```

### Shared Tooltip

`shared = TRUE` shows a single tooltip listing all mark values at the same x
position. Useful for comparing multiple series.

```{r}
df = data.frame(
  x = rep(1:5, 2), y = c(3, 1, 4, 1, 5, 2, 7, 1, 8, 3),
  group = rep(c('A', 'B'), each = 5)
)
p2 = g2(df, y ~ x, color = ~ group) |> mark_line()
p2 |> tooltip_(shared = TRUE)
```

### Group Name

`groupName = FALSE` hides the series/group name prefix in each tooltip item.

```{r}
p2 |> tooltip_(shared = TRUE, groupName = FALSE)
```

## Mark-Level Tooltip (`mark_*(tooltip = list(...))`)

Pass a `tooltip` argument to any `mark_*()` call to control what data is shown
in the tooltip for that specific mark.

| Option | Description |
|--------|-------------|
| `channel` | Channel name to show, e.g. `'y'` or `'x'` |
| `field` | Raw data field name, e.g. `'Temp'` |
| `name` | Custom label for the tooltip item |
| `valueFormatter` | [d3-format](https://d3js.org/d3-format) string, e.g. `'.1f'`, `'.0%'` |
| `color` | Custom color swatch for the item |
| `title` | Title spec: `list(channel = 'x')` or `list(field = 'Date')` |
| `items` | List of item specs for full control over multiple items |
| `FALSE` | Disable tooltip for this mark entirely |

### Show a Single Channel

Use `channel` to restrict the tooltip to one data channel. Hover over any bar
to see only the temperature value instead of both month and temperature.

```{r}
p_bar = g2(air_monthly, Temp ~ Month)
p_bar |> mark_interval(tooltip = list(channel = 'y'))
```

### Custom Item Name

Use `name` to replace the default field name with a descriptive label.

```{r}
p_bar |> mark_interval(tooltip = list(channel = 'y', name = 'Avg Temp (°F)'))
```

### Value Formatter

Use `valueFormatter` with a d3-format string to format the displayed value.

```{r}
p_bar |> mark_interval(tooltip = list(channel = 'y', name = 'Avg Temp', valueFormatter = '.1f'))
```

```{r}
# Percentage formatter in a normalized stacked bar chart
wb = aggregate(breaks ~ wool + tension, data = warpbreaks, FUN = mean)
g2(wb, breaks ~ tension, color = ~ wool) |>
  mark_interval(tooltip = list(channel = 'y', valueFormatter = '.2%')) |>
  transform_('stackY') |>
  transform_('normalizeY') |>
  axis_y(labelFormatter = '.0%') |>
  legend_color(position = 'right') |>
  title_('Warp Breaks by Tension and Wool Type')
```

### Custom Item Color

Use `color` to override the color swatch shown next to the tooltip item.

```{r}
p_bar |> mark_interval(tooltip = list(channel = 'y', name = 'Avg Temp', color = 'tomato'))
```

### Custom Title

Use `title` with a field reference to display a specific data field as the
tooltip title. Use `items` for full control over multiple tooltip entries.

```{r}
air_daily = airquality[, c('Month', 'Day', 'Temp')]
air_daily$Date = as.character(
  as.Date(paste('1973', air_daily$Month, air_daily$Day, sep = '-'))
)
p_daily = g2(air_daily, Temp ~ Date) |> mark_line() |> mark_point()
p_daily |> mark_point(
  tooltip = list(
    title = list(channel = 'x'),
    items = list(list(channel = 'y', name = 'Temp (°F)'))
  )
)
```

### Disable Tooltip for One Mark

Set `tooltip = FALSE` on a specific mark to suppress its tooltip entry while
keeping it for others. In this dual-axis chart, the bars have no tooltip; only
the wind speed line shows a value on hover.

```{r}
air_tw = aggregate(cbind(Temp, Wind) ~ Month, data = airquality, FUN = mean)
air_tw$Month = month.abb[air_tw$Month]
g2(air_tw, ~ Month) |>
  mark_interval(
    encode = list(y = 'Temp'), tooltip = FALSE,
    style = list(fill = '#91caff', fillOpacity = 0.8)
  ) |>
  scale_y(independent = TRUE) |>
  axis_y(title = 'Temp (°F)') |>
  mark_line(
    encode = list(y = 'Wind'),
    style = list(stroke = 'tomato', lineWidth = 2)
  ) |>
  scale_y(independent = TRUE) |>
  axis_y(position = 'right', title = 'Wind (mph)')
```
